### 2016-02-24 writing/hard_stack_paper/paper.tex
Initial outline of the paper.

\documentclass[11pt,letterpaper]{article}
\usepackage{../acl2015}
\usepackage{times}
\usepackage{latexsym}
% \setlength\titlebox{5cm}    % Expanding the titlebox

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% %%%

\title{NSHRDLU:\thanks{Name is provisional, and stolen from Geoff Hinton.}\\Work in progress on the joint learning of parsing and semantic encoding}

\author{
Samuel R.\ Bowman$^{\ast\dag}$ \\
\texttt{sbowman@stanford.edu} \\
\And
Jon Gauthier$^{\dag\ddag}$ \\
\texttt{angeli@stanford.edu} \\
\AND
Christopher D.\ Manning$^{\ast\dag\S}$\\
\texttt{manning@stanford.edu}\\
\And
Christopher Potts$^{\ast}$\\
\texttt{cgpotts@stanford.edu}
\AND\\[-3ex]
{$^{\ast}$Stanford Linguistics\quad
$^{\dag}$Stanford NLP Group}\\
{$^{\ddag}$Stanford Symbolic Systems\quad
$^{\S}$Stanford Computer Science}
}

@@

\begin{document}
\maketitle

@@

This project aims to use ideas from greedy transition-based parsing to build neural network models that can jointly learn to parse sentences and to use those parses to guide semantic composition.

Table~\ref{models-table} shows the sequence of model designs that we plan to build, and Figure~\ref{m1-views} depicts some representative models.

We see three reasons to pursue this approach:
\begin{itemize}
\item Simply adapting a greedy transition-based approach to sentence encoding makes it possible to exploit semantic compositionality in the same manner as in a TreeRNN, but using a static graph structure that can take advantage of existing neural network libraries like Theano for both automatic differentiation and for highly optimized matrix computations on both CPUs and GPUs. Model 0 pursues this property directly, and all subsequent models share it.
\item When a parsing system and an interpretation system are trained jointly and forced to share representations, it is likely that the performance of both models will benefit: the semantic composition function will have better access to syntactic type information that could provide additional evidence on the functional behavior of rare words, and the syntactic parser will have access to information about the semantic interpretability of constituents, making semantically-conditioned parsing decisions like PP attachment easier to learn. Models 1--3 have this property.
\item When a system that uses constituent structure information as a latent variable is initialized from scratch and trained solely to perform some semantic task, it will likely learn some coherent syntax for the genre of language on which it is trained. If this syntax is reasonably stable across similar training sets and across random initializations of training, it can reasonably be said to be latent in the text itself, offering a new kind of evidence about natural language syntax. Model 4 aims to collect this kind of evidence.

\end{itemize}

All five models function as sentence encoding models: their inputs are sentences (input as sequences of words, with the help of a learned embedding matrix), and their outputs are single sentence encoding vectors which can be used as the inputs to downstream models for tasks like sentiment analysis, translation, or inference. Provided that these downstream models are differentiable---as is the case for neural network models and simpler regression models---their gradient signals can be used to train the sentence encoding models.

\begin{table*}[t]
  \centering\small
  \begin{tabular}{cccccc} 
    \toprule
    Name & Stack Representation & Input Representation & Ops Classifier & \multicolumn{2}{l}{Op Predictions Used In} \\ 
      &   &  &   & Training & Testing \\ 
    \midrule
Model 0 & Discrete & Op. sequence & N & -- & -- \\
Model 1 & Discrete & \bf Discrete Buffer & \bf Y: Directly supervised & \bf N & Y \\
Model 2 & Discrete & Discrete Buffer & Y: Directly supervised & \bf Y & Y \\
Model 2S & Discrete & Discrete Buffer & Y: Directly supervised & \bf Sched. sampling & Y \\
Model 3 & \bf Soft & \bf Soft Buffer & Y: Directly supervised & Y & Y \\
Model 4 & Soft & Soft Buffer & \bf Y: Indirectly supervised & Y & Y \\
    \bottomrule
  \end{tabular}
  \protect\caption{\protect\label{models-table}Model variants, ordered by increasing reliance on learning. Bolding indicates the differences between each model and its parent model.} 
\end{table*}

\subsection{Publication plans}

\subsubsection{Winter 2016}

Our first paper will be focused on Model 0 and Model 2S, with no discussion of soft stack models. Our primary goal with this paper will be to achieve state-of-the-art performance on SNLI with Model 0, and performance with Model 2S that surpasses that of a baseline LSTM encoder. We will pursue the state of the art both in the Siamese architecture regime with sentence encoders and in the \citealt{rocktaschel2015reasoning}-style word-by-word attention regime (reimplemented as constituent-by-constituent attention (See \S\ref{sec:c-by-c}).

\subsubsection{Spring 2015}

Our second paper will be focused on the soft stack models, Models 3 and 4. It will investigate the role of differentiable stack data structures in sentence encoding and the possibility of learning consistent and useful syntactic parsers using only semantic supervision.

\section{Models}

@@

\subsection{Model 3}

Model 3 modifies Model 2 by introducing the soft stack/soft queue from \cite{grefenstette2015learning} in place of the hard, conventional stack and buffer. The soft stack makes it possible to for the model to predict smooth distributions over operations of the form (0.93 \shift, 0.07 \reduce), instead of making hard decisions. These soft decisions allow for gradient information to flow from the stack and the buffer back into the operation classifier (blue). This is crucial to our ultimate goal, as it makes it possible for semantic considerations to influence the model's parsing decisions.

Model 3 still receives a direct supervision signal from some existing parser. In order to train the soft stack, we must represent the hard supervision signal from the parser by a soft prediction which matches the soft stack operation output. The supervision signal simply assigns 100\% weight to the ground-truth operation.

\subsection{Model 4}

Model 4 modifies Model 3 by removing the direct supervision signal from the operation classifier (blue), instead forcing the operation classifier to learn solely from the gradient provided by the downstream supervision task. It may be possible to accelerate or otherwise improve Model 4 training by initializing it with parameters learned by Model 3.

By removing an external parser signal, we allow Model 4 to fully exploit the soft stack representation. It is free to predict soft parse operations in the case of ambiguous parses. This distinguishes it from Model 3, which is encouraged to replicate the 100\%-certain ground truth parse predictions provided by the external parser.

\section{Other possible model features}

\subsection{Contextually-informed composition}

The composition function in the basic model (green) combines only the top elements of the stack, without using any further information. We can encourage the composition function to learn to do some amount of context-sensitive interpretation/disambiguation by adding a connection from the tracking LSTM (red) directly into the composition function.

For Model 0, no tracking LSTM is needed for the ordinary operation of the model, but we can simply add one for this purpose, taking as inputs the top two values of the stack at each time point and emitting as output a context vector that can be used to condition the composition function.

So far, we have found this technique to yield non-trivial performance gain. A clearer evaluation on converged models is forthcoming.

\subsection{Constituent-by-constituent attention}\label{sec:c-by-c}

We aim to build on the results of \citealt{rocktaschel2015reasoning} and \citealt{wang2015learning}, who find that neural soft attention models is an extremely effective technique for learning natural language inference. In particular, both papers use versions of word-by-word entailment, in which a latent alignment is learned from every word in the hypothesis to one or more words of the premise. We propose to borrow this basic idea, but to adapt it to a tree-structured setting, proposing \textit{constituent-by-constituent} attention. While these models do attention over a matrix $\mathbf{Y}$ of word-in-context representations from the premise encoder, we will perform attention instead over our own primary data structure, $\mathbf{Y}^{st}$, the matrix of vectors that have appeared at the top of the stack during premise encoding, which correspond one-to-one to the constituents in the tree structure representing the premise. Similarly, while the previous models perform one instance of soft attention conditioned on each word in the hypothesis, we perform one instance of soft attention conditioned on each stack top in the hypothesis encoder, representing the constituents of the hypothesis tree.

\todo{[Anyone -- Sam can] Draw a diagram indicating the structure of constituent by constituent attention.}

\subsection{Encoding the contents of the stack and buffer}

\note{Not currently planned.}

The tracking LSTM (red) needs access to the top of the buffer and the top two elements of the stack in order to make even minimally informed decisions about whether to shift or reduce. It could benefit further from additional information about broader sentential context. This can be provided by running new LSTMs along the elements of each of the stack and the buffer (following \citealt{dyer-EtAl:2015:ACL-IJCNLP}) and feeding the result into the tracking LSTM.

\subsection{Typed \reduce~operations}

\note{Not currently planned.}

Shift-reduce parsers for natural language typically operate with a restricted set of typed \reduce~operations (also known as ``arc'' operations). These operations specify the precise relation between the elements being merged. It would be possible to train any of the parse-supervised models (1--3) to learn such typed arc operations, expanding the op set dramatically to something like \{\shift, \reduce-NP, \reduce-S, \reduce-PP, ...\} (in the case of constituency parse supervision). The model can then learn a distinct composition function depending on the relation of the two elements being merged.

\section{Implementation notes}

The size of the stack should be $N$ for sentences of $N$ words, in case the first reduce merges the final two words. The size of the buffer should be $N - 2$.

@@

\subsection{Memory management and the thin stack}

\todo{[JG] Explain.}

@@

\subsection{Sentiment}

\note{We're running sentiment experiments to evaluate the behavior of converged models, since SNLI takes too long to train to convergence. However, we don't currently plan to publish SNLI results.}

Learning to jointly parse and to predict logical relations between sentences over SST  
\cite{socher2013recursive}.

Evaluation metrics: accuracy, F1 (for all models but 0).

\subsection{Natural language inference}

Learning to jointly parse and to predict logical relations between sentences over SNLI  \cite{snli:emnlp2015}.

Evaluation metrics: accuracy, F1 (for all models but 0).

\subsection{Parser quality evaluation}

\todo{[RG] Write up the parser experiment results.}

@@

\subsection{Inferred Model 4 parses?}

>>>>

\documentclass[11pt,letterpaper]{article}
\usepackage{../acl2015}
\usepackage{times}
\usepackage{latexsym}
% \setlength\titlebox{7.5cm}    % Expanding the titlebox

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Tree-structured neural networks revisited}

\author{
Samuel R.\ Bowman$^{\ast\dag}$ \\
\texttt{sbowman@stanford.edu} \\
\And
Jon Gauthier$^{\dag\ddag}$ \\
\texttt{jgauthie@stanford.edu} \\
\And
Abhinav Rastogi$^{\dag\S}$ \\
\texttt{arastogi@stanford.edu} \\
\AND
Raghav Gupta$^{\dag\mathparagraph}$ \\
\texttt{rgupta93@stanford.edu} \\
\And
Christopher D.\ Manning$^{\ast\dag\mathparagraph}$\\
\texttt{manning@stanford.edu}\\
\And
Christopher Potts$^{\ast}$\\
\texttt{cgpotts@stanford.edu}
\AND\\[-3ex]
{$^{\ast}$Stanford Linguistics\quad
$^{\dag}$Stanford NLP Group\quad
$^{\ddag}$Stanford Symbolic Systems}\\
{$^{\S}$Stanford Electrical Engineering\quad
$^{\mathparagraph}$Stanford Computer Science}
}

@@

\begin{document}
\maketitle
\todo{Anonymize.}

\begin{abstract}
\begin{itemize}
\item TreeRNNs are great, don't work for big data.
\item We present the transition-based TreeRNN implementation -- makes efficient training possible, show that it helps relative to seq models.
\item Further, attention has proven powerful for extending sequence models. We introduce constit-by-constit attention, and find it to be effective here, too.
\end{itemize}
\end{abstract}

@@

\begin{itemize}
\item Brief history of TreeRNNs
\item Speed issues
\item Attention issues
\end{itemize}

\section{The transition-based sentence model}

\subsection{Intuition: Transition-based parsing}

\subsection{The model}

\todo{[SB] Reframe in terms of Model 0 alone.}

@@

\subsection{Tracking left context with an auxiliary LSTM} Includes both the tracking LSTM and context-sensitive composition.

\subsection{Building a model}

\paragraph{The size of the stack}
The size of the stack should be $N$ for sentences of $N$ words, in case the first reduce merges the final two words. The size of the buffer should be $N$.

\paragraph{Handling overly long sentences}

@@

\paragraph{Representing the stack in memory} Or, thin stack. 
\todo{[JG] Explain.}

\paragraph{Optimization and hyperparameters}

\paragraph{Software infrastructure} We will make Theano code available.

@@

\section{Attention over trees}

\paragraph{Constituent-by-constituent attention}
We aim to build on the results of \citealt{rocktaschel2015reasoning} and \citealt{wang2015learning}, who find that neural soft attention models is an extremely effective technique for learning natural language inference. In particular, both papers use versions of word-by-word entailment, in which a latent alignment is learned from every word in the hypothesis to one or more words of the premise. We propose to borrow this basic idea, but to adapt it to a tree-structured setting, proposing \textit{constituent-by-constituent} attention. While these models do attention over a matrix $\mathbf{Y}$ of word-in-context representations from the premise encoder, we will perform attention instead over our own primary data structure, $\mathbf{Y}^{st}$, the matrix of vectors that have appeared at the top of the stack during premise encoding, which correspond one-to-one to the constituents in the tree structure representing the premise. Similarly, while the previous models perform one instance of soft attention conditioned on each word in the hypothesis, we perform one instance of soft attention conditioned on each stack top in the hypothesis encoder, representing the constituents of the hypothesis tree.

\subsection{Accumulating the results of attention}

\paragraph{Accumulating results sequentially} (After Rocktäschel, Wang and Jiang)

\paragraph{Accumulating results using a second tree layer}

\paragraph{The Stanford Natural Language Inference task} \cite{snli:emnlp2015}

\paragraph{Results}

@@

\paragraph{Patterns in the examples that the tree model got right that the LSTM didn't}

\paragraph{Induced alignments}

\section{Conclusions and future work}

\begin{itemize}
\item Future work: Incorporate the parser
\item Future work: Latent parses
\end{itemize}

### 2016-02-24 writing/tree_attn_fig.tex
Initial outline of the paper.

%!TEX root = gist.tex

\begin{figure*}[t]
\centering
\scalebox{0.8}{
 \begin{tikzpicture}
    \tikzstyle{word}=[fill=black!20,text height=2mm]    
    \tikzstyle{nonleaf}=[fill=black!20,text height=2mm]

@@

    \tikzstyle{attn} = [line width=1pt,draw=black!40,opacity=0.25,line cap=round]
    \tikzstyle{heavy} = [line width=2pt]
    \tikzstyle{light} = [line width=0.5pt]
    \tikzstyle{focus} = [draw=blue!40,opacity=1.0]

@@

 \caption{Soft alignments between the nodes of two trees, with alignments to ``the kitten'' highlighted in blue.}\label{fig:tree_attn}
\end{figure*}

>>>>

%!TEX root = hard_stack_paper/paper.tex

\begin{figure}[t]
\centering
\scalebox{0.9}{
 \begin{tikzpicture}
    \tikzstyle{word}=[fill=yellow!40,text height=2mm]    
    \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm]

@@

    \tikzstyle{attn} = [line width=1pt,draw=black!80,opacity=0.25,line cap=round]
    \tikzstyle{heavy} = [line width=3pt]
    \tikzstyle{light} = [line width=0.33pt]
    \tikzstyle{focus} = [draw=red,opacity=1.0]

@@

 \caption{Soft alignments between the nodes of two trees, with alignments to ``the kitten'' highlighted in red.}\label{fig:tree_attn}
\end{figure}
